---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#library("readxl")
#data <- read_excel("D:\\DSSC\\1semester-Courses\\SMDS\\final project\\train.data.xlsx")

data <- read.csv("C:/Users/Flavia/Documents/DSSC/Statistical methods/Exam/mydata3_FINALE.csv")
data$Date <- as.Date(data$Date)


data$log_Recovered_today <- log(data$Recovered_today)
#data$log_Test_today <- log(data$Test_today)
#data$Test_today_1000 <- data$Test_today/1000
data$log10_Variation_cases <- log10(data$Variation_cases)


d.train <- data[1:135 ,] #13th Jan 
d.train

#to check the plot without outliers
d.train.new <- d.train[-c(93,116,128),]

d.test <- data[136:145 ,] #14th to 23th jan
d.test


```

```{r}

library(PerformanceAnalytics)


chart.Correlation(d.train[ ,c("Intensive_care", "Hos_symp","Total_Hos","Home_con",
                              "Total_pos")])

# removing Total_Hos- Total_pos
chart.Correlation(d.train[ ,c("Intensive_care","Hos_symp","Home_con",
                          "Variation_pos","Variation_cases","Recovered_today","Death_today","Test_today",
                          "People_today")])


# removing Home_con
chart.Correlation(d.train[ ,c("Intensive_care","Hos_symp",
                          "Variation_pos","Variation_cases","Recovered_today","Death_today","Test_today",
                          "People_today")])


#removing Variation_pos - People_today
chart.Correlation(d.train[ ,c("Intensive_care","Hos_symp",
                           "Variation_cases","Recovered_today","Death_today","Test_today")])



#removing Variation_pos - Hos_symp- People_today
#this variables are used to build the model:
#chart.Correlation(d.train[ ,c("Intensive_care","Variation_cases",
                             # "Recovered_today","Death_today","Test_today")])


#checking the relations between removed variables

plot(d.train$Hos_symp, d.train$Total_Hos, col=4, lwd=2, xlab="Hospitalised with symptoms",
     ylab="Total hospitalised")

plot(d.train$Home_con, d.train$Hos_symp, col=4, lwd=2, xlab="Home confinement", ylab="Hospitalised with symptoms")

plot(d.train$Total_pos, d.train$Hos_symp, col=4, lwd=2, xlab="Total positives", ylab="Hospitalised with symptoms")


plot(d.train$People_today, d.train$Test_today, col=4, lwd=2, xlab="People_today", ylab="Test_today" )



plot(d.train$People_today, d.train$Intensive_care, col=2, lwd=2, xlab="People_today", ylab="Intensive_care" )

plot(d.train$Variation_pos, d.train$Intensive_care, col=2, lwd=2, xlab="Variation_pos", ylab="Intensive_care")
plot(d.train.new$Variation_pos, d.train.new$Intensive_care, col=2, lwd=2, xlab="Variation_pos", ylab="Intensive_care", sub = "After removing 3 outliers")

```
``` {r}
#relationship between response variable and covariates


#Date
plot(d.train$Date, d.train$Intensive_care , col=2, lwd=2)

#Hos_symp
plot(d.train$Hos_symp, d.train$Intensive_care , col=2, lwd=2)

#Variation_cases
plot(d.train$Variation_cases, d.train$Intensive_care , col=2, lwd=2)
plot(d.train$log10_Variation_cases, d.train$Intensive_care , col=2, lwd=2)

library(ggplot2)
ggplot(d.train, aes(Variation_cases,Intensive_care ))+
  geom_point()+
  stat_smooth(method= glm, formula= y~ poly(x,2, raw=TRUE))

ggplot(d.train, aes(Variation_cases,Intensive_care ))+
  geom_point()+
  stat_smooth(method= glm, formula= y~ poly(x,3, raw=TRUE))


ggplot(d.train, aes(Variation_cases,Intensive_care ))+
  geom_point()+
  stat_smooth(method= glm, formula= y~ poly(x,4, raw=TRUE))


ggplot(d.train, aes(log10_Variation_cases,Intensive_care ))+
  geom_point()+
  stat_smooth(method= glm, formula= y~ poly(x,2, raw=TRUE))



#Recovered_today
plot(d.train$Recovered_today, d.train$Intensive_care , col=2, lwd=2)
plot(d.train.new$Recovered_today, d.train.new$Intensive_care , col=2, lwd=2, sub ="After removing 3 outliers" )

#Death_today
plot(d.train$Death_today, d.train$Intensive_care , col=2, lwd=2)
ggplot(d.train, aes(Death_today,Intensive_care ))+
  geom_point()+
  stat_smooth(method= glm, formula= y~ poly(x,3, raw=TRUE))

ggplot(d.train, aes(Death_today,Intensive_care ))+
  geom_point()+
  stat_smooth(method= glm, formula= y~ poly(x,2, raw=TRUE))


#Test_today
plot(d.train$Test_today, d.train$Intensive_care , col=2, lwd=2)
```

```{r}
#all the dataset plots

#Date
plot(data$Date, data$Intensive_care , col=2, lwd=2, main = " train and test dataset")

#Hos_symp
plot(data$Hos_symp, data$Intensive_care , col=2, lwd=2, main = " train and test dataset")

#Variation_cases
plot(data$Date, data$Variation_cases, col=2, lwd=2, main = " train and test dataset")
plot(d.train$log10_Variation_cases, d.train$Intensive_care , col=2, lwd=2)

library(ggplot2)
ggplot(data, aes(Variation_cases,Intensive_care ))+
  geom_point()+
  stat_smooth(method= glm, formula= y~ poly(x,2, raw=TRUE))

ggplot(data, aes(Variation_cases,Intensive_care ))+
  geom_point()+
  stat_smooth(method= glm, formula= y~ poly(x,3, raw=TRUE))

#Recovered_today
plot(data$Recovered_today, data$Intensive_care , col=2, lwd=2, main = " train and test dataset")
plot(data$log_Recovered_today, data$Intensive_care , col=2, lwd=2, main = " train and test dataset")

#Death_today
plot(data$Death_today, data$Intensive_care , col=2, lwd=2, main = " train and test dataset")
ggplot(data, aes(Death_today,Intensive_care ))+
  geom_point()+
  stat_smooth(method= glm, formula= y~ poly(x,3, raw=TRUE))

ggplot(data, aes(Death_today,Intensive_care ))+
  geom_point()+
  stat_smooth(method= glm, formula= y~ poly(x,2, raw=TRUE))

#Test_today
plot(data$Test_today, data$Intensive_care , col=2, lwd=2, main = " train and test dataset")

```
```{r}
#baseline model

glm0 <- glm(Intensive_care ~ Date+poly(Variation_cases,2)+Hos_symp, 
            family = poisson, data=d.train)

summary(glm0)
AIC <- round(AIC(glm0),2)
BIC <- round(BIC(glm0),2)
Residual_deviance <- round(glm0$deviance,2)

P_value <- pchisq(glm0$deviance,glm0$df.residual, lower.tail = F)

data.frame(AIC, BIC, Residual_deviance,P_value)

library(DAAG)
vif(glm0)

```

we add `Death_today`

```{r}
#adding Death_today

model.glm <- glm(Intensive_care ~ Date+poly(Variation_cases,2)+Hos_symp+
                   poly(Death_today,3), family = poisson, data=d.train)
summary(model.glm)
AIC <- round(AIC(model.glm),2)
BIC <- round(BIC(model.glm),2)
Residual_deviance <- round(model.glm$deviance,2)

#P_value <- pchisq(model.glm$deviance,model.glm$df.residual, lower.tail = F)

data.frame(AIC, BIC, Residual_deviance)

vif(model.glm)

```
we dont add `Recovered_today` due to negligible improvement in AIC
```{r}
#adding Recovered_today

model.glm <- glm(Intensive_care ~ Date+poly(Variation_cases,2)+Hos_symp+ 
                   poly(Death_today,3)+Recovered_today, family = poisson, data=d.train)
summary(model.glm)
AIC <- round(AIC(model.glm),2)
BIC <- round(BIC(model.glm),2)
Residual_deviance <- round(model.glm$deviance,2)

#P_value <- pchisq(model.glm$deviance,model.glm$df.residual, lower.tail = F)

data.frame(AIC, BIC, Residual_deviance)

vif(model.glm)
```


we dont add `Test_today` due to VIF greater than 10 for some variables.
```{r}
# adding Test_today

model.glm <- glm(Intensive_care ~ Date+poly(Variation_cases,2)+Hos_symp+ 
                   poly(Death_today,3)+Test_today, family = poisson, data=d.train)

summary(model.glm)
AIC <- round(AIC(model.glm),2)
BIC <- round(BIC(model.glm),2)
Residual_deviance <- round(model.glm$deviance,2)

#P_value <- pchisq(model.glm$deviance,model.glm$df.residual, lower.tail = F)

data.frame(AIC, BIC, Residual_deviance)

vif(model.glm)
```
we also tried the reverse process:first adding Test then adding Recovered, the result was the same both of them should not be included in the model.


``` {r}
#glm model

model.glm <- glm(Intensive_care ~ Date+poly(Variation_cases,2)+Hos_symp+ 
                   poly(Death_today,3), family = poisson, data=d.train)

summary(model.glm)
AIC <- round(AIC(model.glm),2)
BIC <- round(BIC(model.glm),2)
Residual_deviance <- round(model.glm$deviance,2)

P_value <- pchisq(model.glm$deviance,model.glm$df.residual, lower.tail = F)

data.frame(AIC, BIC, Residual_deviance, P_value)

vif(model.glm)

# #pearson residuals
# p.glm <- residuals(model.glm , type="pearson")
# plot(model.glm$fitted.values, p.glm)


plot(model.glm, which = c(5))
#standardized residuals
z.glm <- (d.train$Intensive_care- model.glm$fitted.values)/sqrt(model.glm$fitted.values)
plot(model.glm$fitted.values,z.glm, xlab = "Fitted values", ylab = "Std. residuals", main = "Standardized residuals- GLM model")
abline(h=-1.96, col="red")
abline(h=1.96, col="red")

par(mfrow=c(2,2))
plot(model.glm, which = c(1,3,2,4))



```

The above glm model is overdispersed. So we try other models:
``` {r}
# quasi piosson

model.glm.quasi <- glm(Intensive_care ~ Date+poly(Variation_cases,2)+Hos_symp+ 
                   poly(Death_today,3), family = quasipoisson, data=d.train)

summary(model.glm.quasi)


#pearson residuals
# p.glm.quasi <- residuals(model.glm.quasi , type="pearson")
# plot(model.glm.quasi$fitted.values, p.glm.quasi)


plot(model.glm.quasi, which = c(5))
#standardized residuals
z.glm.quasi <- (d.train$Intensive_care-model.glm.quasi$fitted.values)/
  sqrt(summary(model.glm.quasi)$dispersion * model.glm.quasi$fitted.values)

plot(model.glm.quasi$fitted.values,z.glm.quasi, xlab = "Fitted values", ylab = "std. residuals", main = "Standardized residuals- Quasipoisson model", ylim=c(-3,4))
abline(h=-1.96, col="red")
abline(h=1.96, col="red")

par(mfrow=c(2,2))
plot(model.glm.quasi, which = c(1,3,2,4))


```



```{r}
# nb
library(MASS)

model.glm.nb <- glm.nb(Intensive_care ~ Date+poly(Variation_cases,2)+Hos_symp+
                   poly(Death_today,3), data=d.train)

summary(model.glm.nb)


#pearson residuals
 p.glm.nb <- residuals(model.glm.nb , type="pearson")
 plot(model.glm.nb$fitted.values, p.glm.nb, main="Pearson residuals- NB model")

#standardized residuals
z.glm.nb <- (d.train$Intensive_care- model.glm.nb$fitted.values)/
  sqrt(model.glm.nb$fitted.values+ (model.glm.nb$fitted.values)^2/summary(model.glm.nb)$dispersion)
plot(model.glm.nb$fitted.values,z.glm.nb, xlab = "Fitted values", ylab = "std. residuals", main = "Standardized residuals- NB model")
abline(h=-1.96, col="red")
abline(h=1.96, col="red")


plot(model.glm.nb, which = c(5))

par(mfrow=c(2,2))
plot(model.glm.nb, which = c(1,3,2,4))

plot(d.train$Variation_cases, d.train$Intensive_care, xlab = "Variation_cases", ylab="Fitted values")
points(d.train$Variation_cases, model.glm$fitted.values,col="blue", pch=19,bg="blue")
points(d.train$Variation_cases, model.glm.nb$fitted.values, col="green", pch=20, bg="green")
legend("bottomright", c("GLM","NB"), pch=c(19,20), col=c("blue", "green"))

plot(d.train$Death_today, d.train$Intensive_care, xlab = "Death_today", ylab="Fitted values")
points(d.train$Death_today, model.glm$fitted.values,col="blue", pch=19,bg="blue")
points(d.train$Death_today, model.glm.nb$fitted.values, col="green", pch=20, bg="green")
legend("bottomright", c("GLM","NB"), pch=c(19,20), col=c("blue", "green"))

plot(d.train$Hos_symp, d.train$Intensive_care, xlab = "Hos_symp", ylab="Fitted values")
points(d.train$Hos_symp, model.glm$fitted.values,col="blue", pch=19,bg="blue")
points(d.train$Hos_symp, model.glm.nb$fitted.values, col="green", pch=20, bg="green")
legend("bottomright", c("GLM","NB"), pch=c(19,20), col=c("blue", "green"))

```
we dont need smooth function for Death_today in gam model.
``` {r}
#gam0
library(mgcv)

gam0 <- gam(Intensive_care ~ Date+s(Variation_cases)+ Hos_symp+s(Death_today)
                   , family = poisson, data=d.train)

summary(gam0)
gam0$sp

plot(gam0, residuals = T, pch=20)
```

``` {r}
# improved gam
library(mgcv)

model.gam <- gam(Intensive_care ~ Date+s(Variation_cases)+ Hos_symp+Death_today
                   , family = poisson, data=d.train)
summary(model.gam)
model.gam$sp

plot(model.gam, residuals = T, pch=20)

plot(d.train$Variation_cases, d.train$Intensive_care, xlab = "Variation_cases", ylab="Fitted values")
points(d.train$Variation_cases, model.glm$fitted.values,col="blue", pch=19,bg="blue")
points(d.train$Variation_cases, model.gam$fitted.values, col="red", pch=20, bg="red")
legend("bottomright", c("GLM","GAM"), pch=c(19,20), col=c("blue", "red"))

plot(d.train$Death_today, d.train$Intensive_care, xlab = "Death_today", ylab="Fitted values")
points(d.train$Death_today, model.glm$fitted.values,col="blue", pch=19,bg="blue")
points(d.train$Death_today, model.gam$fitted.values, col="red", pch=20, bg="red")
legend("bottomright", c("GLM","GAM"), pch=c(19,20), col=c("blue", "red"))

plot(d.train$Hos_symp, d.train$Intensive_care, xlab = "Hos_symp", ylab="Fitted values")
points(d.train$Hos_symp, model.glm$fitted.values,col="blue", pch=19,bg="blue")
points(d.train$Hos_symp, model.gam$fitted.values, col="red", pch=20, bg="red")
legend("bottomright", c("GLM","GAM"), pch=c(19,20), col=c("blue", "red"))

```


```{r}
#random forest
library(randomForest)
model.rf <- randomForest(Intensive_care ~ Date+Variation_cases+Hos_symp+Death_today,
                         data=d.train, importance=TRUE)
model.rf
model.rf$importance

```
It seems that GAM is the best model.
``` {r}
#comparing final models
aic <- AIC(model.glm,model.glm.quasi, model.glm.nb, model.gam)
bic <- BIC(model.glm,model.glm.quasi, model.glm.nb, model.gam)
library(dplyr)
bic2 <- select(bic, BIC)

cbind.data.frame(aic, bic2)

```
``` {r}
#commenting the estimates of gam
# 
# exp(model.gam$coefficients %*% c(1,1, mean(d.train$Variation_cases)+100, (mean(d.train$Variation_cases)+100)^2, mean(d.train$Death_today)+(mean(d.train$Death_today))^2+ (mean(d.train$Death_today))^3+ mean(d.train$Test_today))-
#       exp(model.nb$coefficients %*% c(1,1, mean(d.train$Variation_cases), (mean(d.train$Variation_cases))^2, mean(d.train$Death_today)+(mean(d.train$Death_today))^2+ (mean(d.train$Death_today))^3+ mean(d.train$Test_today))
# 
# invlogit(fit.1$coefficients%*%c(1,mean(data$dist)+1))-
#  invlogit(fit.1$coefficients%*%c(1,mean(data$dist)))
# 
# 
# 
# plot(d.train$Variation_cases, d.train$Intensive_care, ylim = c(100,500),
#      xlab="Variation_cases", ylab="The mean number of Intensive care)",
#      type="n", xaxs="i", yaxs="i", mgp=c(2,.5,0))
# 
# curve(exp(cbind(1, 1,x,x^2,50,50^2,50^3,1000) %*% coef(model.glm.nb)), lwd=.5, add=TRUE)
# curve(exp(cbind(1, 1,x,x^2,50,50^2,50^3,3000) %*% coef(model.glm.nb)), lwd=.5, add=TRUE)
# 
# points(d.train$Variation_cases, d.train$Intensive_care, pch=20, cex=.1)
# text (50, .27, "if As = 0.5", adj=0, cex=.8)
# text (75, .50, "if As = 1.0", adj=0, cex=.8)

```

``` {r}
#predictions

pred.glm <- predict(model.glm, d.test, type= "response")
pred.glm.quasi <- predict(model.glm.quasi, d.test, type= "response")
pred.glm.nb <- predict(model.glm.nb, d.test, type= "response")
pred.gam <- predict(model.gam, d.test, type= "response")
pred.rf <- predict(model.rf, d.test, type = "response")
cbind.data.frame(d.test$Intensive_care,pred.glm,pred.glm.quasi,pred.glm.nb,pred.gam,pred.rf )

#Mean Squre Error
MSE.glm <- mean((d.test$Intensive_care-pred.glm)^2)
MSE.glm.quasi <- mean((d.test$Intensive_care-pred.glm.quasi)^2)
MSE.glm.nb <- mean((d.test$Intensive_care-pred.glm.nb)^2)
MSE.gam <- mean((d.test$Intensive_care-pred.gam)^2)
MSE.rf <- mean((d.test$Intensive_care-pred.rf)^2)
#RMSE
RMSE.glm <- sqrt(MSE.glm)
RMSE.glm.quasi <- sqrt(MSE.glm.quasi)
RMSE.glm.nb <- sqrt(MSE.glm.nb)
RMSE.gam <- sqrt(MSE.gam)
RMSE.rf <- sqrt(MSE.rf)
#NRMSE
NRMSE.glm <- (RMSE.glm)/mean(d.test$Intensive_care)
NRMSE.glm.quasi <- (RMSE.glm.quasi)/mean(d.test$Intensive_care)
NRMSE.glm.nb <- (RMSE.glm.nb)/mean(d.test$Intensive_care)
NRMSE.gam <- (RMSE.gam)/mean(d.test$Intensive_care)
NRMSE.rf <- (RMSE.rf)/mean(d.test$Intensive_care)

models <- c("glm.poisson","glm.quasipoisson","glm.NB","GAM","RF")
MSE <- c(MSE.glm,MSE.glm.quasi,MSE.glm.nb,MSE.gam,MSE.rf)
RMSE <- c(RMSE.glm,RMSE.glm.quasi,RMSE.glm.nb,RMSE.gam,RMSE.rf)
NRMSE <- c(NRMSE.glm,NRMSE.glm.quasi,NRMSE.glm.nb,NRMSE.gam,NRMSE.rf)

data.frame(models,MSE, RMSE,NRMSE)

#library(SimDesign)
#RMSE(d.test$Intensive, pred.glm, type="CV")

library(ggplot2)
 ggplot()+ 
   ggtitle("Predictions from 14th to 23th Jan. 2021")+
    geom_line(data=d.test, aes(x=Date, y=Intensive_care, color="actual"), size=1)+
    geom_point(data = d.test,aes(x=Date, y=Intensive_care ))+
    geom_line(data=d.test, aes(x=Date, y=pred.glm, color="Poisson"), size=1)+
    geom_line(data=d.test, aes(x=Date, y=pred.glm.quasi, color="Quasi Poisson"), size=1)+
    geom_line(data=d.test, aes(x=Date, y=pred.glm.nb, color="Negative Binomial"), size=1)+
    geom_line(data=d.test, aes(x=Date, y=pred.gam, color="GAM"), size=1)+
    geom_line(data=d.test, aes(x=Date, y=pred.rf, color= "RF"), size=1)+
  geom_point()+
   scale_color_manual(name= "models", values = c("actual"="green","Poisson"="red", "Quasi Poisson"= "blue", "Negative Binomial"= "orange", "GAM"= "#FF9999" , "RF"="purple"))+
   xlab("Date") + ylab ("Intensive care") 

# poisson and quasi poisson have the same coefficients. so the same plots
```



``` {r}
#adding new variables
plot(d.train$Date, d.train$Zone_color,  col=2, lwd=2)
plot(d.train$Date, d.train$lag_zone_color,  col=2, lwd=2)
plot(d.train$Zone_color, d.train$Intensive_care, col=2, lwd=2)
plot(d.train$lag_zone_color, d.train$Intensive_care, col=2, lwd=2)

```

According to gam model, we dont consider a polynomial for `Death_today` in our models.
```{r}
#new glm (linear Death_today)
model.glm1 <- glm(Intensive_care ~ Date+poly(Variation_cases,2)+Hos_symp+
                   Death_today, family = poisson, data=d.train)
summary(model.glm1)
aic <- AIC(model.glm1)
bic <- BIC(model.glm1)
RD <- model.glm1$deviance

data.frame(aic, bic, RD)

vif(model.glm1)
```

we add `Zone_color`
```{r}
#adding Zone_color
model.glm1 <- glm(Intensive_care ~ Date+poly(Variation_cases,2)+Hos_symp+
                   Death_today+Zone_color, family = poisson, data=d.train)
summary(model.glm1)
aic <- AIC(model.glm1)
bic <- BIC(model.glm1)
RD <- model.glm1$deviance

data.frame(aic, bic, RD)

vif(model.glm1)

```

we add `lag_zone_color`
```{r}
#adding lag_zone_color
model.glm1 <- glm(Intensive_care ~ Date+poly(Variation_cases,2)+Hos_symp+
                   Death_today+Zone_color+lag_zone_color, family = poisson, data=d.train)
summary(model.glm1)
aic <- AIC(model.glm1)
bic <- BIC(model.glm1)
RD <- model.glm1$deviance

data.frame(aic, bic, RD)

vif(model.glm1)

```
we add season
```{r}
#adding season
model.glm1 <- glm(Intensive_care ~ Date+poly(Variation_cases,2)+Hos_symp+
                   Death_today+Zone_color+lag_zone_color+Season, family = poisson, data=d.train)
summary(model.glm1)
aic <- AIC(model.glm1)
bic <- BIC(model.glm1)
RD <- model.glm1$deviance

data.frame(aic, bic, RD)

vif(model.glm1)

```
we dont add any interaction.
```{r}
#adding interaction
model.glm1 <- glm(Intensive_care ~ Date+poly(Variation_cases,2)+Hos_symp+
                   Death_today+Zone_color+lag_zone_color+Season+Season:Hos_symp, family = poisson, data=d.train)
summary(model.glm1)
aic <- AIC(model.glm1)
bic <- BIC(model.glm1)
RD <- model.glm1$deviance

data.frame(aic, bic, RD)

vif(model.glm1)

```

``` {r}
#new glm 
model.glm1 <- glm(Intensive_care ~ Date+poly(Variation_cases,2)+Hos_symp+
                   Death_today+Zone_color+lag_zone_color+Season, family = poisson, data=d.train)
summary(model.glm1)
aic <- AIC(model.glm1)
bic <- BIC(model.glm1)
RD <- model.glm1$deviance
p_val <- pchisq(model.glm1$deviance, model.glm1$df.residual, lower.tail = F)

data.frame(aic, bic, RD, p_val)

vif(model.glm1)

#par(mfrow=c(2,2))
plot(model.glm1, which = c(1,2,3,4,5))

# #pearson residuals
# p.glm <- residuals(model.glm , type="pearson")
# plot(model.glm$fitted.values, p.glm)

#standardized residuals
z.glm <- (d.train$Intensive_care- model.glm1$fitted.values)/sqrt(model.glm1$fitted.values)
plot(model.glm1$fitted.values,z.glm, xlab = "Fitted values", ylab = "std. residuals", main = "Standardized residuals- GLM model")
abline(h=-1.96, col="red")
abline(h=1.96, col="red")

```
```{r}
# new quasi piosson
model.glm.quasi1 <- glm(Intensive_care ~ Date+poly(Variation_cases,2)+Hos_symp+
                   Death_today+Zone_color+lag_zone_color+Season, family = quasipoisson, data=d.train)
summary(model.glm.quasi1)



#par(mfrow=c(2,2))
plot(model.glm.quasi1, which = c(1,2,3,4,5))

#pearson residuals
# p.glm.quasi <- residuals(model.glm.quasi , type="pearson")
# plot(model.glm.quasi$fitted.values, p.glm.quasi)

#standardized residuals
z.glm.quasi <- (d.train$Intensive_care-model.glm.quasi1$fitted.values)/
  sqrt(summary(model.glm.quasi1)$dispersion * model.glm.quasi1$fitted.values)

plot(model.glm.quasi1$fitted.values,z.glm.quasi, xlab = "Fitted values", ylab = "std. residuals", main = "Standardized residuals- Quasipoisson model", ylim=c(-3,4))
abline(h=-1.96, col="red")
abline(h=1.96, col="red")

```

```{r}
# nb
library(MASS)
model.glm.nb1 <- glm.nb(Intensive_care ~ Date+poly(Variation_cases,2)+Hos_symp+
                   Death_today+Zone_color+lag_zone_color+Season, 
                   data=d.train)
summary(model.glm.nb1)



#par(mfrow=c(2,2))
plot(model.glm.nb1, which = c(1,2,3,4,5))

#pearson residuals
 p.glm.nb <- residuals(model.glm.nb1 , type="pearson")
 plot(model.glm.nb1$fitted.values, p.glm.nb, main="Pearson residuals- NB model")

#standardized residuals
z.glm.nb <- (d.train$Intensive_care- model.glm.nb1$fitted.values)/
  sqrt(model.glm.nb1$fitted.values+ (model.glm.nb1$fitted.values)^2/summary(model.glm.nb1)$dispersion)
plot(model.glm.nb1$fitted.values,z.glm.nb, xlab = "Fitted values", ylab = "std. residuals", main = "Standardized residuals- NB model")
abline(h=-1.96, col="red")
abline(h=1.96, col="red")


plot(d.train$Variation_cases, d.train$Intensive_care, xlab = "Variation_cases", ylab="Fitted values")
points(d.train$Variation_cases, model.glm1$fitted.values,col="blue", pch=19,bg="blue")
points(d.train$Variation_cases, model.glm.nb1$fitted.values, col="green", pch=20, bg="green")
legend("bottomright", c("GLM","NB"), pch=c(19,20), col=c("blue", "green"))

plot(d.train$Death_today, d.train$Intensive_care, xlab = "Death_today", ylab="Fitted values")
points(d.train$Death_today, model.glm1$fitted.values,col="blue", pch=19,bg="blue")
points(d.train$Death_today, model.glm.nb1$fitted.values, col="green", pch=20, bg="green")
legend("bottomright", c("GLM","NB"), pch=c(19,20), col=c("blue", "green"))

plot(d.train$Hos_symp, d.train$Intensive_care, xlab = "Hos_symp", ylab="Fitted values")
points(d.train$Hos_symp, model.glm1$fitted.values,col="blue", pch=19,bg="blue")
points(d.train$Hos_symp, model.glm.nb1$fitted.values, col="green", pch=20, bg="green")
legend("bottomright", c("GLM","NB"), pch=c(19,20), col=c("blue", "green"))

```
``` {r}
#gam
library(mgcv)

model.gam1 <- gam(Intensive_care ~ Date+s(Variation_cases)+Hos_symp+
                   Death_today+Zone_color+lag_zone_color+Season
                   , family = poisson, data=d.train)
summary(model.gam1)
model.gam1$sp

plot(model.gam1, residuals = T, pch=20)

plot(d.train$Variation_cases, d.train$Intensive_care, xlab = "Variation_cases", ylab="Fitted values")
points(d.train$Variation_cases, model.glm1$fitted.values,col="blue", pch=19,bg="blue")
points(d.train$Variation_cases, model.gam1$fitted.values, col="red", pch=20, bg="red")
legend("bottomright", c("GLM","GAM"), pch=c(19,20), col=c("blue", "red"))

plot(d.train$Death_today, d.train$Intensive_care, xlab = "Death_today", ylab="Fitted values")
points(d.train$Death_today, model.glm1$fitted.values,col="blue", pch=19,bg="blue")
points(d.train$Death_today, model.gam1$fitted.values, col="red", pch=20, bg="red")
legend("bottomright", c("GLM","GAM"), pch=c(19,20), col=c("blue", "red"))

plot(d.train$Hos_symp, d.train$Intensive_care, xlab = "Hos_symp", ylab="Fitted values")
points(d.train$Hos_symp, model.glm1$fitted.values,col="blue", pch=19,bg="blue")
points(d.train$Hos_symp, model.gam1$fitted.values, col="red", pch=20, bg="red")
legend("bottomright", c("GLM","GAM"), pch=c(19,20), col=c("blue", "red"))

```
```{r}
#random forest
library(randomForest)
model.rf1 <- randomForest(Intensive_care ~ Date+Variation_cases+Hos_symp+
                   Death_today+Zone_color+lag_zone_color+Season, data=d.train, importance=TRUE)
model.rf1
model.rf1$importance

```

``` {r}
#comparing final models
aic <- AIC(model.glm1,model.glm.quasi1, model.glm.nb1, model.gam1)
bic <- BIC(model.glm1,model.glm.quasi1, model.glm.nb1, model.gam1)

library(dplyr)
bic2 <- select(bic, BIC)

cbind.data.frame(aic, bic2)
```
again gam is the best model.  

``` {r}
#predictions

pred.glm1 <- predict(model.glm1, d.test, type= "response")
pred.glm.quasi1 <- predict(model.glm.quasi1, d.test, type= "response")
pred.glm.nb1 <- predict(model.glm.nb1, d.test, type= "response")
pred.gam1 <- predict(model.gam1, d.test, type= "response")
pred.rf1 <- predict(model.rf1, d.test, type = "response")
cbind.data.frame(d.test$Intensive_care,pred.glm1,pred.glm.quasi1,pred.glm.nb1,pred.gam1,pred.rf1 )


#Mean Squre Error
MSE.glm1 <- mean((d.test$Intensive_care-pred.glm1)^2)
MSE.glm.quasi1 <- mean((d.test$Intensive_care-pred.glm.quasi1)^2)
MSE.glm.nb1 <- mean((d.test$Intensive_care-pred.glm.nb1)^2)
MSE.gam1 <- mean((d.test$Intensive_care-pred.gam1)^2)
MSE.rf1 <- mean((d.test$Intensive_care-pred.rf1)^2)
#RMSE
RMSE.glm1 <- sqrt(MSE.glm1)
RMSE.glm.quasi1 <- sqrt(MSE.glm.quasi1)
RMSE.glm.nb1 <- sqrt(MSE.glm.nb1)
RMSE.gam1 <- sqrt(MSE.gam1)
RMSE.rf1 <- sqrt(MSE.rf1)
#NRMSE
NRMSE.glm1 <- (RMSE.glm1)/mean(d.test$Intensive_care)
NRMSE.glm.quasi1 <- (RMSE.glm.quasi1)/mean(d.test$Intensive_care)
NRMSE.glm.nb1 <- (RMSE.glm.nb1)/mean(d.test$Intensive_care)
NRMSE.gam1 <- (RMSE.gam1)/mean(d.test$Intensive_care)
NRMSE.rf1 <- (RMSE.rf1)/mean(d.test$Intensive_care)

models <- c("glm.poisson","glm.quasipoisson","glm.NB","GAM","RF")
MSE <- c(MSE.glm1,MSE.glm.quasi1,MSE.glm.nb1,MSE.gam1,MSE.rf1)
RMSE <- c(RMSE.glm1,RMSE.glm.quasi1,RMSE.glm.nb1,RMSE.gam1,RMSE.rf1)
NRMSE <- c(NRMSE.glm1,NRMSE.glm.quasi1,NRMSE.glm.nb1,NRMSE.gam1,NRMSE.rf1)

data.frame(models,MSE, RMSE,NRMSE)

library(ggplot2)
 ggplot()+ 
   ggtitle("Predictions from 14th to 23th Jan. 2021")+
    geom_line(data=d.test, aes(x=Date, y=Intensive_care, color="actual"), size=1)+
    geom_point(data = d.test,aes(x=Date, y=Intensive_care ))+
    geom_line(data=d.test, aes(x=Date, y=pred.glm1, color="Poisson"), size=1)+
    geom_line(data=d.test, aes(x=Date, y=pred.glm.quasi1, color="Quasi Poisson"), size=1)+
    geom_line(data=d.test, aes(x=Date, y=pred.glm.nb1, color="Negative Binomial"), size=1)+
    geom_line(data=d.test, aes(x=Date, y=pred.gam1, color="GAM"), size=1)+
    geom_line(data=d.test, aes(x=Date, y=pred.rf1, color= "RF"), size=1)+
  
   scale_color_manual(name= "models", values = c("actual"="green","Poisson"="red", "Quasi Poisson"= "blue", "Negative Binomial"= "orange", "GAM"= "#FF9999" , "RF"="purple"))+
   xlab("Date") + ylab ("Intensive care") 

# poisson and quasi poisson have the same coefficients. so the same plots
```

